---
title: "R Notebook"
output: html_notebook
---

```{r}
locked_sens <- function(spec,sens) {
  tfun <- approxfun(spec,sens,ties="ordered")
  lsens <- tfun(0.99)
  return(lsens)
}

calc_sens <- function(data,size,seed) {
  set.seed(size)
  sens_vec <- numeric(100)
  for (i in 1:100) {
  print(i)
  class_index5 <- createDataPartition(data$tumor, p = .8, list = FALSE)
  class_train5 <- data[class_index5, ]
  class_test5 <- data[-class_index5, ]
  # slnrf_fit5 <- train(as.factor(tumor) ~ .,
  #                     data = class_train5[],
  #                     method = "lda",  ## can also do "lasso, "ridge", ...etc
  #                     preProc = c("center","scale")) 
  rnrf_fit7 <- train(as.factor(tumor) ~ .,
                     data = class_train5[],
                     method = "rf",  ## can also do "lasso, "ridge", ...etc
                     preProc = c("center","scale")) 
  # gnrf_fit7 <- train(as.factor(tumor) ~ .,
  #                    data = class_train5[],
  #                    method = "glm",  ## can also do "lasso, "ridge", ...etc
  #                    preProc = c("center","scale")) 
  # nrf_fit5 <- train(as.factor(tumor) ~ .,
  #                   data = class_train5[],
  #                   method = "naive_bayes",  ## can also do "lasso, "ridge", ...etc
  #                   preProc = c("center","scale")) 
  # slnrf5_roc <- predict(slnrf_fit5, class_test5, type="prob")
  # slnrf5_tree <- roc(as.character(class_test5$tumor), slnrf5_roc[,"0"])
  # r <- predict(rnrf_fit7, class_test5)
  rnrf7_roc <- predict(rnrf_fit7, class_test5, type="prob")
  # cf7b <- confusionMatrix(r, as.factor(class_test5$tumor))
  rnrf7_tree <- roc(as.character(class_test5$tumor), rnrf7_roc[,as.character(size)])
  # g <-  predict(gnrf_fit7, class_test5)
  # gnrf7_roc <- predict(gnrf_fit7, class_test5, type="prob")
  # cf7b2 <- confusionMatrix(g, as.factor(class_test5$tumor))
  # gnrf7_tree <- roc(as.character(class_test5$tumor), gnrf7_roc[,"7"])
  # nrf5_roc <- predict(nrf_fit5, class_test5, type="prob")
  # nrf5_tree <- roc(as.character(class_test5$tumor), nrf5_roc[,"0"])
  # ldasens2[i] <- slnrf5_tree$sensitivities[slnrf5_tree$specificities >= 0.99][1]
  # ldaauc2[i] <- slnrf5_tree$auc
  sens_vec[i] <- locked_sens(rnrf7_tree$specificities,rnrf7_tree$sensitivities)
  # rfsens7b2[i] <- mean(rnrf7_tree$sensitivities[rnrf7_tree$specificities >= 0.99])
  # rfauc7b[i] <- rnrf7_tree$auc
  # rfspec7b[i] <- cf7b$byClass[2]
  # glmsens7b[i] <-  locked_sens(gnrf7_tree$specificities,gnrf7_tree$sensitivities)
  # glmsens7b2[i] <- mean(gnrf7_tree$sensitivities[gnrf7_tree$specificities >= 0.99])
  # glmauc7b[i] <- gnrf7_tree$auc
  # glmspec7b[i] <- cf7b2$byClass[2]
  # nbsens2[i] <- nrf5_tree$sensitivities[gnrf5_tree$specificities >= 0.99][1]
  # nbauc2[i] <- nrf5_tree$auc
  }
  return(sens_vec)
  
}
```

```{r}
tumor1 <- calc_sens(df1[,-c(1,2,43,44,45)],1,101)
tumor2 <- calc_sens(df2[,-c(1,2,43,44)],2,101)
tumor3 <- calc_sens(df3[,-c(1,2,43,44)],3,101)
tumor4 <- calc_sens(df4[,-c(1,42,43,44,45)],4,101)
tumor5 <- calc_sens(df5[,-c(1,2,43,44)],5,101)
tumor6 <- calc_sens(df6bil,6,101)
tumor7 <- calc_sens(df7[,-c(1,42)],7,101)
tumor8 <- calc_sens(df6[,-c(1,42,43,44,45,46,47,48)],8,101)
```

```{r}
tsize1 <- data.frame(TumorDiameter = 1.2, Method = "CNA",sensitivity = tumor1)
tsize2 <- data.frame(TumorDiameter = 1.6, Method = "CNA",sensitivity = tumor2)
tsize3 <- data.frame(TumorDiameter = 1.8, Method = "CNA",sensitivity = tumor3)
tsize4 <- data.frame(TumorDiameter = 2.0, Method = "CNA",sensitivity = tumor4)
tsize5 <- data.frame(TumorDiameter = 2.1, Method = "CNA",sensitivity = tumor5)
tsize6 <- data.frame(TumorDiameter = 2.3, Method = "CNA",sensitivity = tumor6)
tsize7 <- data.frame(TumorDiameter = 2.4, Method = "CNA",sensitivity = tumor7)
tsize8 <- data.frame(TumorDiameter = 2.5, Method = "CNA",sensitivity = tumor8)
msize1 <- data.frame(TumorDiameter = 1.2, Method = "Mutation",sensitivity = 0.048)
msize2 <- data.frame(TumorDiameter = 1.6, Method = "Mutation",sensitivity = 0.272)
msize3 <- data.frame(TumorDiameter = 1.8, Method = "Mutation",sensitivity = 0.601)
msize4 <- data.frame(TumorDiameter = 2.0, Method = "Mutation",sensitivity = 0.842)
msize5 <- data.frame(TumorDiameter = 2.1, Method = "Mutation",sensitivity = 0.953)
msize6 <- data.frame(TumorDiameter = 2.3, Method = "Mutation",sensitivity = 0.989)
msize7 <- data.frame(TumorDiameter = 2.4, Method = "Mutation",sensitivity = 0.998)
msize8 <- data.frame(TumorDiameter = 2.5, Method = "Mutation",sensitivity = 1.000)
tumor_sizes <- rbind(tsize1,tsize2,tsize3,tsize4,tsize5,tsize6,tsize7,tsize8,msize1,msize2,msize3,
                     msize4,msize5,msize6,msize7,msize8)
library(dplyr)
tumor_sizes_agg <- tumor_sizes %>%
  group_by(TumorDiameter,Method) %>%
  summarise( 
    n=n(),
    mean=mean(sensitivity),
    sd=sd(sensitivity)
  ) %>%
  mutate( se=sd/sqrt(n))  %>%
  mutate( ic=se * qt((1-0.05)/2 + .5, n-1))
figure3.1 <- ggplot(tumor_sizes_agg,aes(x=TumorDiameter,y=mean,color=Method)) +
  geom_line() + geom_point() +
  geom_errorbar( aes(x=TumorDiameter, ymin=mean-ic, ymax=mean+ic), width=0.1, colour="orange", alpha=0.9, size=0.7) +
  ggtitle("Tumor Detection using CNA vs Mutations") + labs(x = "Tumor Diameter (cm)", y = "Average Sensitivity at 99% Specificity") + theme(text=element_text(size=13))
```


100 Train-Test Split WGS for Line Plot
```{r}
library(caret)
library(pROC)
###### 1 bil
# glmsens1b <- numeric(100)
# glmsens1b2 <- numeric(100)
# glmauc1b <- numeric(100)
# rfsens1b <- numeric(100)
# rfsens1b2 <- numeric(100)
# rfauc1b <- numeric(100)
# glmspec1b <- numeric(100)
# rfspec1b <- numeric(100)
# BAFWGS1bil <- read_csv("~/Downloads/BAFWGS1bil.csv")
# df1 <- data.frame(BAFWGS1bil)
# df1.1 <- df1[,-c(1,2,43)]
# set.seed(101)
#### 2 bil
# glmsens2b <- numeric(100)
# glmsens2b2 <- numeric(100)
# glmauc2b <- numeric(100)
# rfsens2b <- numeric(100)
# rfsens2b2 <- numeric(100)
# rfauc2b <- numeric(100)
# glmspec2b <- numeric(100)
# rfspec2b <- numeric(100)
# BAFWGS2bil <- read_csv("~/Downloads/BAFWGS2bil.csv")
# df2 <- data.frame(BAFWGS2bil)
# df2.1 <- df2[,-c(1,2,43)]
##### 3 bil
#BAFWGS5bil <- read_csv("~/Downloads/BAFWGS5bil.csv")
#df5 <- data.frame(BAFWGS5bil)
# df3.1 <- df3[,-c(1,2,43)]
# glmsens3b <- numeric(100)
# glmsens3b2 <- numeric(100)
# glmauc3b <- numeric(100)
# rfsens3b <- numeric(100)
# rfsens3b2 <- numeric(100)
# rfauc3b <- numeric(100)
# glmspec3b <- numeric(100)
# rfspec3b <- numeric(100)
##### 4 bil
# glmsens4b <- numeric(100)
# glmsens4b2 <- numeric(100)
# glmauc4b <- numeric(100)
# rfsens4b <- numeric(100)
# rfsens4b2 <- numeric(100)
# rfauc4b <- numeric(100)
# glmspec4b <- numeric(100)
# rfspec4b <- numeric(100)
# df4.11 <- df4.1[,-c(1,42,43,44,45)]
# set.seed(101)
###### 6 bil
# glmsens6b <- numeric(100)
# glmsens6b2 <- numeric(100)
# glmauc6b <- numeric(100)
# rfsens6b <- numeric(100)
# rfsens6b2 <- numeric(100)
# rfauc6b <- numeric(100)
# glmspec6b <- numeric(100)
# rfspec6b <- numeric(100)
# BAFWGS6bil <- read_csv("~/Downloads/BAFWGS6bil.csv")
# df6bil <- data.frame(BAFWGS6bil)
# df6bil <- df6bil[,-c(1,2,43)]
######## 7 bil
glmsens7b <- numeric(100)
# glmsens7b2 <- numeric(100)
# glmauc7b <- numeric(100)
rfsens7b <- numeric(100)
# rfsens7b2 <- numeric(100)
# rfauc7b <- numeric(100)
# glmspec7b <- numeric(100)
# rfspec7b <- numeric(100)
BAFWGS7bil <- read_csv("~/Downloads/BAFWGS7bil.csv")
df7 <- data.frame(BAFWGS7bil)
df7 <- df7[,-c(1,2,43)]
set.seed(101)
# Add error bars for points
# Do 1-10 billion cells for plot
for (i in 1:100) {
  print(i)
  class_index5 <- createDataPartition(df7$tumor, p = .8, list = FALSE)
  class_train5 <- df7[class_index5, ]
  class_test5 <- df7[-class_index5, ]
  # slnrf_fit5 <- train(as.factor(tumor) ~ .,
  #                     data = class_train5[],
  #                     method = "lda",  ## can also do "lasso, "ridge", ...etc
  #                     preProc = c("center","scale")) 
  rnrf_fit7 <- train(as.factor(tumor) ~ .,
                     data = class_train5[],
                     method = "rf",  ## can also do "lasso, "ridge", ...etc
                     preProc = c("center","scale")) 
  gnrf_fit7 <- train(as.factor(tumor) ~ .,
                     data = class_train5[],
                     method = "glm",  ## can also do "lasso, "ridge", ...etc
                     preProc = c("center","scale")) 
  # nrf_fit5 <- train(as.factor(tumor) ~ .,
  #                   data = class_train5[],
  #                   method = "naive_bayes",  ## can also do "lasso, "ridge", ...etc
  #                   preProc = c("center","scale")) 
  # slnrf5_roc <- predict(slnrf_fit5, class_test5, type="prob")
  # slnrf5_tree <- roc(as.character(class_test5$tumor), slnrf5_roc[,"0"])
  r <- predict(rnrf_fit7, class_test5)
  rnrf7_roc <- predict(rnrf_fit7, class_test5, type="prob")
  cf7b <- confusionMatrix(r, as.factor(class_test5$tumor))
  rnrf7_tree <- roc(as.character(class_test5$tumor), rnrf7_roc[,"7"])
  g <-  predict(gnrf_fit7, class_test5)
  gnrf7_roc <- predict(gnrf_fit7, class_test5, type="prob")
  cf7b2 <- confusionMatrix(g, as.factor(class_test5$tumor))
  gnrf7_tree <- roc(as.character(class_test5$tumor), gnrf7_roc[,"7"])
  # nrf5_roc <- predict(nrf_fit5, class_test5, type="prob")
  # nrf5_tree <- roc(as.character(class_test5$tumor), nrf5_roc[,"0"])
  # ldasens2[i] <- slnrf5_tree$sensitivities[slnrf5_tree$specificities >= 0.99][1]
  # ldaauc2[i] <- slnrf5_tree$auc
  rfsens7b[i] <- locked_sens(rnrf7_tree$specificities,rnrf7_tree$sensitivities)
  # rfsens7b2[i] <- mean(rnrf7_tree$sensitivities[rnrf7_tree$specificities >= 0.99])
  # rfauc7b[i] <- rnrf7_tree$auc
  # rfspec7b[i] <- cf7b$byClass[2]
  glmsens7b[i] <-  locked_sens(gnrf7_tree$specificities,gnrf7_tree$sensitivities)
  # glmsens7b2[i] <- mean(gnrf7_tree$sensitivities[gnrf7_tree$specificities >= 0.99])
  # glmauc7b[i] <- gnrf7_tree$auc
  # glmspec7b[i] <- cf7b2$byClass[2]
  # nbsens2[i] <- nrf5_tree$sensitivities[gnrf5_tree$specificities >= 0.99][1]
  # nbauc2[i] <- nrf5_tree$auc
}
```

100 Train-Test Split WES Testing 4 Different Models
```{r}
# myfunc <- func(sens,spec) {} returns sensitivity at 99%
BAFWES4bil <- read_csv("~/Downloads/BAFWES4bil.csv")
edf4 <- data.frame(BAFWES4bil)
edf4.1 <- edf4[,-c(1,2,43)]
eldasens1 <- numeric(100)
eldaauc1 <- numeric(100)
erfsens1 <- numeric(100)
erfauc1 <- numeric(100)
eglmsens1 <- numeric(100)
eglmauc1 <- numeric(100)
enbsens1 <- numeric(100)
enbauc1 <- numeric(100)
set.seed(101)
for (i in 1:100) {
  print(i)
  eclass_index5 <- createDataPartition(edf4.1$tumor, p = .8, list = FALSE)
  eclass_train5 <- edf4.1[eclass_index5, ]
  eclass_test5 <- edf4.1[-eclass_index5, ]
  eslnrf_fit5 <- train(as.factor(tumor) ~ .,
                      data = eclass_train5[],
                      method = "lda",  ## can also do "lasso, "ridge", ...etc
                      preProc = c("center","scale")) 
  ernrf_fit5 <- train(as.factor(tumor) ~ .,
                     data = eclass_train5[],
                     method = "rf",  ## can also do "lasso, "ridge", ...etc
                     preProc = c("center","scale")) 
  egnrf_fit5 <- train(as.factor(tumor) ~ .,
                     data = eclass_train5[],
                     method = "glm",  ## can also do "lasso, "ridge", ...etc
                     preProc = c("center","scale")) 
  enrf_fit5 <- train(as.factor(tumor) ~ .,
                    data = eclass_train5[],
                    method = "naive_bayes",  ## can also do "lasso, "ridge", ...etc
                    preProc = c("center","scale")) 
  eslnrf5_roc <- predict(eslnrf_fit5, eclass_test5, type="prob")
  eslnrf5_tree <- roc(as.character(eclass_test5$tumor), eslnrf5_roc[,"4"])
  ernrf5_roc <- predict(ernrf_fit5, eclass_test5, type="prob")
  ernrf5_tree <- roc(as.character(eclass_test5$tumor), ernrf5_roc[,"4"])
  egnrf5_roc <- predict(egnrf_fit5, eclass_test5, type="prob")
  egnrf5_tree <- roc(as.character(eclass_test5$tumor), egnrf5_roc[,"4"])
  enrf5_roc <- predict(enrf_fit5, eclass_test5, type="prob")
  enrf5_tree <- roc(as.character(eclass_test5$tumor), enrf5_roc[,"4"])
  eldasens1[i] <- locked_sens(eslnrf5_tree$specificities,eslnrf5_tree$sensitivities)
  eldaauc1[i] <- eslnrf5_tree$auc
  erfsens1[i] <- locked_sens(ernrf5_tree$specificities,ernrf5_tree$sensitivities)
  erfauc1[i] <- ernrf5_tree$auc
  eglmsens1[i] <- locked_sens(egnrf5_tree$specificities,egnrf5_tree$sensitivities)
  eglmauc1[i] <- egnrf5_tree$auc
  enbsens1[i] <- locked_sens(enrf5_tree$specificities,enrf5_tree$sensitivities)
  enbauc1[i] <- enrf5_tree$auc
}
```

Line Plot for CNA
```{r}
rframe1bil <- melt(data.frame(TumorSize = "1", Method = "CNA", DetectionSensitivity = rfsens1))
rframe2bil <- melt(data.frame(TumorSize = "2", Method = "CNA", DetectionSensitivity = rfsens2b))
rframe3bil <- melt(data.frame(TumorSize = "3", Method = "CNA", DetectionSensitivity = rfsens3b))
rframe4bil <- melt(data.frame(TumorSize = "4", Method = "CNA", DetectionSensitivity = rfsens4b))
rframe5bil <- melt(data.frame(TumorSize = "5", Method = "CNA", DetectionSensitivity = rfsens5b))
rframe6bil <- melt(data.frame(TumorSize = "6", Method = "CNA", DetectionSensitivity = rfsens6b))
rframe7bil <- melt(data.frame(TumorSize = "7", Method = "CNA", DetectionSensitivity = rfsens7b))
rframe8bil <- melt(data.frame(TumorSize = "8", Method = "CNA", DetectionSensitivity = rfsens))
rframe1bilsp <- melt(data.frame(TumorSize = "1", Method = "CNA", DetectionSpecificity = rfspec1b))
rframe2bilsp <- melt(data.frame(TumorSize = "2", Method = "CNA", DetectionSpecificity = rfspec2b))
rframe3bilsp <- melt(data.frame(TumorSize = "3", Method = "CNA", DetectionSpecificity = rfspec3b))
rframe4bilsp <- melt(data.frame(TumorSize = "4", Method = "CNA", DetectionSpecificity = rfspec4b))
rframe5bilsp <- melt(data.frame(TumorSize = "5", Method = "CNA", DetectionSpecificity = rfspec5b))
rframe5bilsp <- melt(data.frame(TumorSize = "5", Method = "CNA", DetectionSpecificity = rfspec5b))
rframe6bilsp <- melt(data.frame(TumorSize = "6", Method = "CNA", DetectionSpecificity = rfspec6b))
rframe7bilsp <- melt(data.frame(TumorSize = "7", Method = "CNA", DetectionSpecificity = rfspec7b))
rframe8bilsp <- melt(data.frame(TumorSize = "8", Method = "CNA", DetectionSpecificity = rfspec))
muframe1bil <- melt(data.frame(TumorSize = "1", Method = "Mutation", DetectionSensitivity = sens1b))
muframe2bil <- melt(data.frame(TumorSize = "2", Method = "Mutation", DetectionSensitivity = sens2b))
muframe3bil <- melt(data.frame(TumorSize = "3", Method = "Mutation", DetectionSensitivity = sens3b))
muframe4bil <- melt(data.frame(TumorSize = "4", Method = "Mutation", DetectionSensitivity = sens4b))
muframe5bil <- melt(data.frame(TumorSize = "5", Method = "Mutation", DetectionSensitivity = sens5b))
muframe6bil <- melt(data.frame(TumorSize = "6", Method = "Mutation", DetectionSensitivity = sens6b))
muframe7bil <- melt(data.frame(TumorSize = "7", Method = "Mutation", DetectionSensitivity = sens7b))
muframe8bil <- melt(data.frame(TumorSize = "8", Method = "Mutation", DetectionSensitivity = sens8b))
# muframe1bilsp <- melt(data.frame(TumorSize = "1", Method = "Mutation", DetectionSpecificity = spec1b))
# muframe2bilsp <- melt(data.frame(TumorSize = "2", Method = "Mutation", DetectionSpecificity = spec2b))
# muframe3bilsp <- melt(data.frame(TumorSize = "3", Method = "Mutation", DetectionSpecificity = spec3b))
# muframe4bilsp <- melt(data.frame(TumorSize = "4", Method = "Mutation", DetectionSpecificity = spec4b))
# muframe5bilsp <- melt(data.frame(TumorSize = "5", Method = "Mutation", DetectionSpecificity = spec5b))
# muframe6bilsp <- melt(data.frame(TumorSize = "6", Method = "Mutation", DetectionSpecificity = spec6b))
# muframe8bilsp <- melt(data.frame(TumorSize = "8", Method = "Mutation", DetectionSpecificity = spec8b))
# lineplot <- rbind(rframe1bil,rframe2bil,rframe3bil,rframe4bil,rframe5bil,rframe6bil,rframe8bil,rframe1bilsp,rframe2bilsp,rframe3bilsp,rframe4bilsp,rframe5bilsp,rframe6bilsp,rframe8bilsp,muframe1bil,muframe2bil,muframe3bil,
#                   muframe4bil,muframe5bil,muframe6bil,muframe8bil,muframe1bilsp,muframe2bilsp,muframe3bilsp,
#                   muframe4bilsp,muframe5bilsp,muframe6bilsp,muframe8bilsp)
lineplot <- rbind(rframe1bil,rframe2bil,rframe3bil,rframe4bil,rframe5bil,rframe6bil,rframe7bil,rframe8bil,rframe1bilsp,rframe2bilsp,rframe3bilsp,rframe4bilsp,rframe5bilsp,rframe6bilsp,rframe7bilsp,rframe8bilsp,muframe1bil,muframe2bil,muframe3bil,
                  muframe4bil,muframe5bil,muframe6bil,muframe7bil,muframe8bil)
lineplot
```

```{r}
# Stack plots vertically, specificity as separate.
library(dplyr)
library(ggplot2)
### Look at package cowplot, plot_grid(plot1,plot2,align=V,col=1) takes two ggplots
lineplot1 <- lineplot %>% group_by(TumorSize,Method,variable) %>% summarise(n=n(),mean=mean(value),sd=sd(value)) %>% mutate(se = sd/sqrt(n)) %>% mutate(ic = se * qt((1-0.05)/2 + 0.5, n-1))
lineplot1$TumorSize <- as.numeric(lineplot1$TumorSize)
lineplotsens <- lineplot1[lineplot1$variable == "DetectionSensitivity",]
lineplotspec <- lineplot1[lineplot1$variable == "DetectionSpecificity",]
lineplotspecCNA <- lineplotspec[lineplotspec$Method == "CNA",]
######
figure3 <- ggplot(lineplotsens,aes(TumorSize,mean,group=Method)) + geom_line(aes(color=Method)) + geom_point(aes(color=Method)) + geom_errorbar(aes(ymin=mean-ic,ymax=mean+ic), width=0.1,color="orange") + geom_line(data=lineplotspecCNA,aes(x=TumorSize,y=mean,group=Method,color="green")) + geom_point(data=lineplotspecCNA,aes(x=TumorSize,y=mean,color="green")) + geom_errorbar(data=lineplotspecCNA,aes(ymin=mean-ic,ymax=mean+ic), width=0.1,color="orange") + labs(x = "Tumor Size (Billions of Cells)", y = "Average Sensitivity")
# figure4 <- ggplot(lineplotspec,aes(TumorSize,mean,group=Method)) + geom_line(aes(color=Method)) + geom_point(aes(color=Method)) + geom_errorbar(aes(ymin=mean-ic,ymax=mean+ic), width=0.1,color="orange")
#lineplot12 <- lineplot1[lineplot1$Method == "Mutation" & lineplot1$variable == "DetectionSpecifcity",]
figure3.1 <- ggplot(lineplot1,aes(x=TumorSize,y=mean,group=Method)) + geom_line(aes(color=Method)) + geom_point(aes(color=Method)) + facet_grid(variable ~ ., scales="free_y") + geom_errorbar(aes(ymin=mean-ic,ymax=mean+ic), width=0.1,color="orange") + labs(x = "Tumor Size (Billions of Cells)", y = "Average Value over 100 Train-Test Split Permutations")
```

Testing Different Classifiers to detect cancer using BAF using WGS
```{r}
# Interpolation function
df4.1 <- df4.1[,-c(1,42,43)]
ldasens1 <- numeric(100)
rfsens1 <- numeric(100)
glmsens1 <- numeric(100)
nbsens1 <- numeric(100)
set.seed(101)
for (i in 1:100) {
  print(i)
  class_index5 <- createDataPartition(df4.1$tumor, p = .8, list = FALSE)
  class_train5 <- df4.1[eclass_index5, ]
  class_test5 <- df4.1[-eclass_index5, ]
  slnrf_fit5 <- train(as.factor(tumor) ~ .,
                      data = class_train5[],
                      method = "lda",  ## can also do "lasso, "ridge", ...etc
                      preProc = c("center","scale")) 
  rnrf_fit5 <- train(as.factor(tumor) ~ .,
                     data = class_train5[],
                     method = "rf",  ## can also do "lasso, "ridge", ...etc
                     preProc = c("center","scale")) 
  gnrf_fit5 <- train(as.factor(tumor) ~ .,
                     data = class_train5[],
                     method = "glm",  ## can also do "lasso, "ridge", ...etc
                     preProc = c("center","scale")) 
  nrf_fit5 <- train(as.factor(tumor) ~ .,
                    data = class_train5[],
                    method = "naive_bayes",  ## can also do "lasso, "ridge", ...etc
                    preProc = c("center","scale")) 
  slnrf5_roc <- predict(slnrf_fit5, class_test5, type="prob")
  slnrf5_tree <- roc(as.character(class_test5$tumor), slnrf5_roc[,"4"])
  rnrf5_roc <- predict(rnrf_fit5, class_test5, type="prob")
  rnrf5_tree <- roc(as.character(class_test5$tumor), rnrf5_roc[,"4"])
  gnrf5_roc <- predict(gnrf_fit5, class_test5, type="prob")
  gnrf5_tree <- roc(as.character(class_test5$tumor), gnrf5_roc[,"4"])
  nrf5_roc <- predict(nrf_fit5, class_test5, type="prob")
  nrf5_tree <- roc(as.character(class_test5$tumor), nrf5_roc[,"4"])
  ldasens1[i] <- locked_sens(slnrf5_tree$specificities,slnrf5_tree$sensitivities)
  rfsens1[i] <- locked_sens(rnrf5_tree$specificities,rnrf5_tree$sensitivities)
  glmsens1[i] <- locked_sens(gnrf5_tree$specificities,gnrf5_tree$sensitivities)
  nbsens1[i] <- locked_sens(nrf5_tree$specificities,nrf5_tree$sensitivities)
}
```

```{r}
# vals <- data.frame("Model" = c("glm","lda","nb","rf"), "AverageSensitivity" = c(mean(glmsens2),mean(ldasens2),mean(nbsens2),mean(rfsens2)))
# figure1.1 <- ggplot(vals,aes(x=Model,y=AverageSensitivity,fill=Model)) + geom_bar(stat="identity") + labs(y = "Average Sensitivity at 99%+ Specificity", title = "Model Testing using WGS + BAF to Detect 4 Billion Cell Tumor (30x Coverage)")
# vals2 <- data.frame("Model" = c("glm","lda","nb","rf"), "AverageSensitivity" = c(mean(eglmsens2),mean(eldasens2),mean(enbsens2),mean(erfsens2)))
# figure2.1 <- ggplot(vals2,aes(x=Model,y=AverageSensitivity,fill=Model)) + geom_bar(stat="identity") + labs(y = "Average Sensitivity at 99%+ Specificity", title = "Model Testing using WES + BAF to Detect 4 Billion Cell Tumor (3000x Coverage)")
nb <- data.frame(model = "nb", sensitivity = nbsens1)
rf <- data.frame(model = "rf", sensitivity = rfsens1)
glm <- data.frame(model = "glm", sensitivity = glmsens1)
lda <- data.frame(model = "lda", sensitivity = ldasens1)
models <- rbind(nb,rf,glm,lda)

nb2 <- data.frame(model = "nb", sensitivity = enbsens1)
rf2 <- data.frame(model = "rf", sensitivity = erfsens1)
glm2 <- data.frame(model = "glm", sensitivity = eglmsens1)
lda2 <- data.frame(model = "lda", sensitivity = eldasens1)
models2 <- rbind(nb2,rf2,glm2,lda2)

library(dplyr)
my_sum <- models %>%
  group_by(model) %>%
  summarise( 
    n=n(),
    mean=mean(sensitivity),
    sd=sd(sensitivity)
  ) %>%
  mutate( se=sd/sqrt(n))  %>%
  mutate( ic=se * qt((1-0.05)/2 + .5, n-1))

my_sum2 <- models2 %>%
  group_by(model) %>%
  summarise( 
    n=n(),
    mean=mean(sensitivity),
    sd=sd(sensitivity)
  ) %>%
  mutate( se=sd/sqrt(n))  %>%
  mutate( ic=se * qt((1-0.05)/2 + .5, n-1))

figure1.1 <- ggplot(my_sum) +
  geom_bar( aes(x=model, y=mean), stat="identity", fill="forestgreen", alpha=0.5) +
  geom_errorbar( aes(x=model, ymin=mean-ic, ymax=mean+ic), width=0.4, colour="orange", alpha=0.9, size=1.5) +
  ggtitle("Model Testing using WGS + BAF to Detect 4 Billion Cell Tumor (30x Coverage)") + labs(y = "Average Sensitivity at 99%+ Specificity")

figure2.1 <- ggplot(my_sum2) +
  geom_bar( aes(x=model, y=mean), stat="identity", fill="forestgreen", alpha=0.5) +
  geom_errorbar( aes(x=model, ymin=mean-ic, ymax=mean+ic), width=0.4, colour="orange", alpha=0.9, size=1.5) +
  ggtitle("Using WES + BAF to Detect 2 cm diameter tumor (3000x Coverage)") + labs(y = "Average Sensitivity at 99% Specificity") + theme(text=element_text(size=12))
```

4 bil WGS plot
```{r}
library(dplyr)
nb1s <- data.frame(model = "nb", sensitivity = nbsens1)
nb1sm <- melt(nb1s)
nbauc <- data.frame(model = "nb", AUC = nbauc2)
nbaucm <- melt(nbauc)
rf1s <- data.frame(model = "rf", sensitivity = rfsens1)
rf1sm <- melt(rf1s)
rfauc <- data.frame(model = "rf", AUC = rfauc2)
rfaucm <- melt(rfauc)
glm1s <- data.frame(model = "glm", sensitivity = glmsens1)
glm1sm <- melt(glm1s)
glmauc <- data.frame(model = "glm", AUC = glmauc2)
glmaucm <- melt(glmauc)
lda1s <- data.frame(model = "lda", sensitivity = ldasens1)
lda1sm <- melt(lda1s)
ldauc <- data.frame(model = "lda", AUC = ldaauc2)
ldaucm <- melt(ldauc)
nb1sp <- data.frame(model = "nb", specificity = test_nbspec)
nb1spm <- melt(nb1sp)
rf1sp <- data.frame(model = "rf", specificity = test_rfspec)
rf1spm <- melt(rf1sp)
glm1sp <- data.frame(model = "glm", specificity = test_glmspec)
glm1spm <- melt(glm1sp)
lda1sp <- data.frame(model = "lda", specificity = test_ldaspec)
lda1spm <- melt(lda1sp)
models1 <- rbind(nb1sm,rf1sm,glm1sm,lda1sm,nb1spm,rf1spm,glm1spm,lda1spm,
                 nbaucm,rfaucm,glmaucm,ldaucm)

# my_sum1 <- models1 %>%
#   group_by(model) %>%
#   summarise( 
#     n=n(),
#     mean_sensitivity=mean(sensitivity),
#     sd_sensitivtiy=sd(sensitivity),
#     mean_specificity=mean(specificity),
#     sd_specificity=sd(specificity),
#     mean_AUC=mean(AUC),
#     sd_AUC=sd(AUC)
#   ) %>%
#   mutate( se_sensitivity=sd_sensitivity/sqrt(n),
#           se_specificity=sd_specificity/sqrt(n),
#           se_AUC=sd_AUC/sqrt(n)) %>%
#   mutate( ic_sensitivity=se_sensitivity * qt((1-0.05)/2 + .5, n-1),
#           ic_specificity=se_specificity * qt((1-0.05)/2 + .5, n-1),
#           ic_AUC=se_AUC * qt((1-0.05)/2 + 0.5, n-1))

# my_sum3 <- models3 %>%
#   group_by(model) %>%
#   summarise( 
#     n=n(),
#     mean=mean(AUC),
#     sd=sd(AUC)
#   ) %>%
#   mutate( se=sd/sqrt(n))  %>%
#   mutate( ic=se * qt((1-0.05)/2 + .5, n-1))

# ggplot(my_sum1) +
#   geom_bar( aes(x=model, y=mean), stat="identity", fill="forestgreen", alpha=0.5) +
#   geom_errorbar( aes(x=model, ymin=mean-ic, ymax=mean+ic), width=0.4, colour="orange", alpha=0.9, size=1.5) +
#   ggtitle("Sensitivity with 95% Confidence Interval") + labs(y = "Sensitivity")
# 
# ggplot(my_sum3) +
#   geom_bar( aes(x=model, y=mean), stat="identity", fill="forestgreen", alpha=0.5) +
#   geom_errorbar( aes(x=model, ymin=mean-ic, ymax=mean+ic), width=0.4, colour="orange", alpha=0.9, size=1.5) +
#   ggtitle("AUC with 95% Confidence Interval") + labs(y = "AUC")
```

```{r}
my_sum1
```

Whole Genome Performance Model Testing
```{r}
my_sum1 <- models1 %>% group_by(model,variable) %>% summarise(n=n(),mean=mean(value),sd=sd(value)) %>% mutate(se = sd/sqrt(n)) %>% mutate(ic = se * qt((1-0.05)/2 + 0.5, n-1))
figure1 <- ggplot(my_sum1) + geom_bar(aes(x=model,y=mean,fill=variable), stat = "identity", position=position_dodge()) + geom_errorbar(aes(x=model,ymin=mean-ic,ymax=mean+ic,fill=variable),position=position_dodge(.9),width=0.4,alpha=0.9,color="black") + labs(title="Model Testing using WGS + BAF to Detect 4 Billion Cell Tumor (30x Coverage)", y="Average Value")
```

4 bil WES plot, Whole Exome Performance Testing
```{r}
# library(dplyr)
# enb1s <- data.frame(model = "nb", sensitivity = enbsens1)
# enb1sm <- melt(enb1s)
# enbauc <- data.frame(model = "nb", AUC = enbauc2)
# enbaucm <- melt(enbauc)
# erf1s <- data.frame(model = "rf", sensitivity = erfsens1)
# erf1sm <- melt(erf1s)
# erfauc <- data.frame(model = "rf", AUC = erfauc2)
# erfaucm <- melt(erfauc)
# eglm1s <- data.frame(model = "glm", sensitivity = eglmsens1)
# eglm1sm <- melt(eglm1s)
# eglmauc <- data.frame(model = "glm", AUC = eglmauc2)
# eglmaucm <- melt(eglmauc)
# elda1s <- data.frame(model = "lda", sensitivity = eldasens1)
# elda1sm <- melt(elda1s)
# eldauc <- data.frame(model = "lda", AUC = eldaauc2)
# eldaucm <- melt(eldauc)
# enb1sp <- data.frame(model = "nb", specificity = etest_nbspec)
# enb1spm <- melt(enb1sp)
# erf1sp <- data.frame(model = "rf", specificity = etest_rfspec)
# erf1spm <- melt(erf1sp)
# eglm1sp <- data.frame(model = "glm", specificity = etest_glmspec)
# eglm1spm <- melt(eglm1sp)
# elda1sp <- data.frame(model = "lda", specificity = etest_ldaspec)
# elda1spm <- melt(elda1sp)
# emodels1 <- rbind(enb1sm,erf1sm,eglm1sm,elda1sm,enb1spm,erf1spm,eglm1spm,elda1spm,
#                  enbaucm,erfaucm,eglmaucm,eldaucm)
# emy_sum1 <- emodels1 %>% group_by(model,variable) %>% summarise(n=n(),mean=mean(value),sd=sd(value)) %>% mutate(se = sd/sqrt(n)) %>% mutate(ic = se * qt((1-0.05)/2 + 0.5, n-1))
# figure2 <- ggplot(emy_sum1) + geom_bar(aes(x=model,y=mean,fill=variable), stat = "identity", position=position_dodge()) + geom_errorbar(aes(x=model,ymin=mean-ic,ymax=mean+ic,fill=variable),position=position_dodge(.9),width=0.4,alpha=0.9,color="black") + labs(title="Model Testing using WES + BAF to Detect 4 Billion Cell Tumor (3000x Coverage)", y="Average Value")
```

-log10 score threshold for 4 billion cells
```{r}
thresh <- seq(0,150,1)
acc <- numeric(151)
sens <- numeric(151)
spec <- numeric(151)
for (i in 1:151) {
  for (j in 1:500) {
    if (df4.1$score[j] >= thresh[i]) {
      df4.1$scoreprediction[j] <- 4
    }
    if (df4.1$score[j] < thresh[i]) {
      df4.1$scoreprediction[j] <- 0
    }
  }
  acc[i] <- length(which(df4.1$tumor == df4.1$scoreprediction)) / 500
  tp <- length(which(df4.1$tumor == 4 & df4.1$scoreprediction == 4))
  tn <- length(which(df4.1$tumor == 0 & df4.1$scoreprediction == 0))
  fn <- length(which(df4.1$tumor == 4 & df4.1$scoreprediction == 0))
  fp <- length(which(df4.1$tumor == 0 & df4.1$scoreprediction == 4))
  sens[i] <- tp / (tp + fn)
  spec[i] <- tn / (tn + fp)
}
thresholds <- data.frame(Thresholds = as.character(thresh), Sensitivity = sens, Specificity = spec)
```

same as above for 1 billion cells
```{r}
thresh1bil <- seq(0,150,1)
acc1bil <- numeric(151)
sens1bil <- numeric(151)
spec1bil <- numeric(151)
for (i in 1:151) {
  for (j in 1:500) {
    if (df1$score[j] >= thresh[i]) {
      df1$scoreprediction[j] <- 1
    }
    if (df1$score[j] < thresh[i]) {
      df1$scoreprediction[j] <- 0
    }
  }
  acc1bil[i] <- length(which(df1$tumor == df1$scoreprediction)) / 500
  tp <- length(which(df1$tumor == 1 & df1$scoreprediction == 1))
  tn <- length(which(df1$tumor == 0 & df1$scoreprediction == 0))
  fn <- length(which(df1$tumor == 1 & df1$scoreprediction == 0))
  fp <- length(which(df1$tumor == 0 & df1$scoreprediction == 1))
  sens1bil[i] <- tp / (tp + fn)
  spec1bil[i] <- tn / (tn + fp)
}
thresholds1bil <- data.frame(Thresholds = as.character(thresh1bil), Accuracy = acc1bil, Sensitivity = sens1bil, Specificity = spec1bil)
mthresholds1bil <- melt(thresholds1bil)
figure6 <- ggplot(mthresholds1bil,aes(x=as.numeric(Thresholds),y=value,group=variable)) + geom_point(aes(color=variable)) + geom_line(aes(color=variable)) + labs(x="Thresholds",title="-Log10 p-value of B-allele frequency (Null = 0.5) as Metric to identify 1 billion cell tumor")
```

```{r}
mthresholds <- melt(thresholds)
figure5 <- ggplot(mthresholds,aes(x=as.numeric(Thresholds),y=value,group=variable)) + geom_point(aes(color=variable)) + geom_line(aes(color=variable)) + labs(x="Thresholds",title="-Log10 p-value of B-allele frequency (Null = 0.5) as Metric to identify 2 cm diameter tumor")
```

```{r}
which(acc == max(acc))
which(acc1bil == max(acc1bil))
```

```{r}
thresh[23]
sens[23]
spec[23]
thresh[15]
sens1bil[15]
spec1bil[15]
```

```{r}
df4
df4.1
##Test df4.1 threshold on df4
```

```{r}
for (j in 1:500) {
    if (df4$score[j] >= 25) {
      df4$scoreprediction[j] <- 4
    }
    if (df4$score[j] < 25) {
      df4$scoreprediction[j] <- 0
    }
}
length(which(df4$tumor == df4$scoreprediction)) / 500
tp <- length(which(df4$tumor == 4 & df4$scoreprediction == 4))
tn <- length(which(df4$tumor == 0 & df4$scoreprediction == 0))
fn <- length(which(df4$tumor == 4 & df4$scoreprediction == 0))
fp <- length(which(df4$tumor == 0 & df4$scoreprediction == 4))
sensthresh <- tp / (tp + fn)
specthresh <- tn / (tn + fp)
sensthresh
specthresh
result <- data.frame("Sensitivity" = sensthresh, "Specificity" = specthresh)
result <- melt(result)
figure7 <- ggplot(result,aes(x=variable,y=value,fill=variable)) + geom_bar(stat="identity") + labs(title="Testing -Log10 p-value of BAF as Threshold to identify 4 billion cell tumor")
# Barplot
```

```{r}
result <- data.frame("Sensitivity" = sensthresh, "Specificity" = specthresh)
result <- melt(result)
figure7 <- ggplot(result,aes(x=variable,y=value,fill=variable)) + geom_bar(stat="identity") + labs(title="Testing -Log10 p-value of BAF as Threshold to identify 2 cm diameter tumor")
```

```{r}
DataFigure2 <- read_csv("~/Downloads/DataFigure2.csv")
```

```{r}
dfig <- data.frame(DataFigure2)
dfig <- dfig[,-c(1)]
library(dplyr)
dfig1 <- dfig %>% group_by(model) %>%
  summarise(n=n(),mean=mean(sensitivity),sd=sd(sensitivity)) %>% mutate(
    se=sd/sqrt(n)) %>% mutate(ic = se * qt((1-0.05)/2 + .5, n-1))

figure1.12 <- ggplot(dfig1) + geom_bar(aes(x=model,y=mean), stat = "identity", fill = "forestgreen", alpha = 0.5) + geom_errorbar(aes(x=model,ymin=mean-ic,ymax=mean+ic), width = 0.4, color = "orange", alpha = 0.9, size = 1.5) + ggtitle("Using WGS + BAF to Detect 2 cm Diameter Tumor (30x Coverage)") + labs(y="Average Sensitivity at 99% Specificity") + theme(text=element_text(size=10))
```

Combinatorial Approach of ML/Random Forest vs crude statistical test (potential value to both)